---
title: "Cross Validation"
output: github_document
---
```{r}
library(tidyverse)
library(modelr)
library(mgcv)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


## Nonlinear data and CV
```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

Do the train / test split

```{r}
train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```


```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

quick visualization of the linear model with smooth mod

```{r}
train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

same visualization of the linear model with wiggly mod
```{r}
train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

comparing the linear models
```{r}
train_df |> 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

compute root mean squared errors (RMSEs) for each model

```{r}
rmse(linear_mod, train_df)
rmse(smooth_mod, train_df)
rmse(wiggly_mod, train_df)

rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```
RMSEs on training data is misleading. While RMSEs on testing data gives a sesne of out-of-sample predicition accuracy.


## Use modelr for CV

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) 
```

```{r}
cv_df |> pull(train) |> nth(1) |> as_tibble()
```

```{r}
cv_df |> pull(test) |> nth(1) |> as_tibble()
```

```{r}
cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

Apply each model to all training datasets, and evaluate on all testing datasets.

```{r}
cv_results = 
  cv_df |> 
  mutate(
    linear_fit = map(train, \(df) lm(y ~ x, data = df)),
    smooth_fit = map(train, \(df) mgcv::gam(y ~ s(x), data = df)),
    wiggly_fit = map(train, \(df) mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = df))
    ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_fit, test, \(mod, df) rmse(mod, df)),
    rmse_smooth = map2_dbl(smooth_fit, test, \(mod, df) rmse(mod, df)),
    rmse_wiggly = map2_dbl(wiggly_fit, test, \(mod, df) rmse(mod, df))
  )
```

```{r}
cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  group_by(model_type) |> 
  summarize(m_rmse = mean(rmse))
```

```{r}
cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  ggplot(aes(x = model_type, y = rmse)) + 
  geom_violin()
```


## Child Growth Example

```{r}
child_growth = read_csv("./data/nepalese_children.csv")
```

```{r}
child_growth |> 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5)
```

```{r}
child_growth =
  child_growth |> 
  mutate(weight_cp = (weight > 7) * (weight - 7))
```

```{r}
linear_mod = lm(armc ~ weight, data = child_growth)
pwl_mod = lm(armc ~ weight + weight_cp, data = child_growth)
smooth_mod = gam(armc ~ s(weight), data = child_growth)
```

```{r}
child_growth |> 
  gather_predictions(linear_mod, pwl_mod, smooth_mod) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

```{r}
cv_df =
  crossv_mc(child_growth, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, ~lm(armc ~ weight, data = .x)),
    pwl_mod = map(train, ~lm(armc ~ weight + weight_cp, data = .x)),
    smooth_mod  = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_pwl = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

 
